---
title: "Mathematical practice final exam 2025"
date: '2025-07-02'
documentclass: ctexart
output:
 pdf_document:
    latex_engine: xelatex
 html_document:
    df_print: paged
---

```{r message = F, echo = F}
# 设置全局代码块选项
knitr::opts_chunk$set(
  echo = TRUE,          # 显示代码
  warning = FALSE,      # 不显示警告信息
  message = FALSE,      # 不显示消息
  fig.align = "center"  # 图片居中显示
)
library(tidyverse)
# install.packages("Ecdat")
library(Ecdat)
library(knitr)  # 用于表格输出
```

## 1. 
Find the inverse of the following matrix and verify it using the `all.equal()` function.
$$
\left(\begin{array}{cccc}
9 & 4 & 12 & 2 \\
5 & 0 & 7 & 9 \\
2 & 6 & 8 & 0 \\
9 & 2 & 9 & 11
\end{array}\right) 
$$
```{r}
# 创建矩阵A
# byrow = FALSE 表示按列填充数据
A <- matrix(c(9, 5, 2, 9, 4, 0, 6, 2, 12, 7, 8, 9, 2, 9, 0, 11),
            nrow = 4, byrow = FALSE)

# 使用 solve() 函数计算矩阵A的逆矩阵
A_inv <- solve(A)

# 打印原始矩阵和其逆矩阵
print("Matrix A:")
kable(A)
print("Inverse of Matrix A:")
kable(A_inv)

# 验证：A乘以其逆矩阵应该得到一个单位矩阵
# 创建一个4x4的单位矩阵用于比较
I <- diag(4)

# 使用 all.equal() 检查 A %*% A_inv 是否约等于单位矩阵 I
# 这个函数可以处理浮点数计算带来的微小误差
verification <- all.equal(A %*% A_inv, I)

# 打印验证结果
print(paste("Verification result (TRUE means success):", verification))
```


## 2. 
Execute the following lines which create two vectors of random integers which are chosen with replacement from the integers $0, 1, \dots , 999$. Both vectors have length 250. 

```{r}
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
n <- length(xVec)
```

(a) Create the vector $(y_2 - x_1, \cdots, y_n - x_{n-1}).$
```{r}
# yVec从第二个元素取到最后一个
# xVec从第一个元素取到倒数第二个
result_a <- yVec[2:n] - xVec[1:(n-1)]
print("First 6 elements of the resulting vector:")
head(result_a)
```


(b) Pick out the values in yVec which are $> 600$.

```{r}
# 使用逻辑索引来筛选yVec中大于600的元素
result_b <- yVec[yVec > 600]
print("Values in yVec > 600 (first 6):")
head(result_b)
```


(c) What are the index positions in yVec of the values which are $> 600$? 

```{r}
# which() 函数返回满足条件的元素的索引
result_c <- which(yVec > 600)
print("Index positions of values > 600 in yVec (first 6):")
head(result_c)
```


(d) Sort the numbers in the vector xVec in the order of increasing values in yVec. 

```{r}
# order(yVec) 返回yVec按升序排列的原始索引
# 然后用这个索引顺序来排列xVec
result_d <- xVec[order(yVec)]
print("xVec sorted by yVec (first 6):")
head(result_d)
```


(e) Pick out the elements in yVec at index positions $1, 4, 7, 10, 13, \cdots$

```{r}
# seq() 生成一个从1开始，步长为3的序列
indices <- seq(from = 1, to = n, by = 3)
result_e <- yVec[indices]
print("Elements from yVec at positions 1, 4, 7, ... (first 6):")
head(result_e)
```

## 3.
For this problem we'll use the (built-in) dataset state.x77.
```{r}
data(state)
state.x77 <- as_tibble(state.x77, rownames  = 'State')

state_data <- state.x77
# 为了方便后续操作，将列名中的空格替换为下划线
colnames(state_data) <- make.names(colnames(state_data))
```

a. Select all the states having an income less than 4300, and calculate the average income of these states.
```{r}
# 使用 filter() 筛选收入低于4300的州
low_income_states <- state_data %>% filter(Income < 4300)

# 使用 summarise() 计算平均收入
avg_income <- low_income_states %>% summarise(AverageIncome = mean(Income))

print("Average income for states with income < 4300:")
kable(avg_income)
```

b.	Sort the data by income and select the state with the highest income.
```{r}
# arrange(desc(Income)) 按收入降序排列
# slice(1) 选取第一行，即收入最高的州
highest_income_state <- state_data %>% 
  arrange(desc(Income)) %>% 
  slice(1)

print("State with the highest income:")
kable(highest_income_state)
```

c.	Add a variable to the data frame which categorizes the size of population: $<= 4500$ is `S`, $> 4500$ is `L`.
```{r}
# 使用 mutate() 和 ifelse() 添加新列 Population.Size
state_data <- state_data %>%
  mutate(
    Population.Size = ifelse(Population <= 4500, "S", "L")
  )

print("Data with new Population.Size column (first 6 rows):")
kable(head(state_data))
```

d.	Find out the average income and illiteracy of the two groups of states, distinguishing by whether the states are small or large.
```{r}
# group_by() 按人口规模分组
# summarise() 分别计算每组的平均收入和平均文盲率
summary_by_size <- state_data %>%
  group_by(Population.Size) %>%
  summarise(
    Average.Income = mean(Income),
    Average.Illiteracy = mean(Illiteracy)
  )

print("Summary by population size:")
kable(summary_by_size)
```

## 4. 
a. Write a function to simulate `n` observations of $(X_1, X_2)$ which follow the uniform distribution over the square $[0, 1] \times [0, 1]$.
```{r}
# 定义函数，输入参数为n (观测值数量)
simulate_points <- function(n) {
  # 使用runif(n)生成n个[0,1]区间的随机数
  tibble(
    X1 = runif(n),
    X2 = runif(n)
  )
}

# 调用函数生成10个点作为示例
example_points <- simulate_points(10)
print("Example of 10 simulated points:")
kable(example_points)
```

b. Write a function to calculate the proportion of the observations that the distance between $(X_1, X_2)$ and the nearest edge is less than 0.25, and the proportion of them with the distance to the nearest vertex less than 0.25.
```{r}
# 定义函数，输入为包含X1, X2坐标的数据框
calculate_proportions <- function(points_df) {
  n_obs <- nrow(points_df)
  
  # 计算每个点到最近边缘的距离
  # pmin()返回每个位置上的最小值
  dist_edge <- pmin(points_df$X1, 1 - points_df$X1, 
                      points_df$X2, 1 - points_df$X2)
  
  # 计算每个点到最近顶点的距离
  # 四个顶点为 (0,0), (1,0), (0,1), (1,1)
  dist_v1 <- sqrt(points_df$X1^2 + points_df$X2^2)
  dist_v2 <- sqrt((1 - points_df$X1)^2 + points_df$X2^2)
  dist_v3 <- sqrt(points_df$X1^2 + (1 - points_df$X2)^2)
  dist_v4 <- sqrt((1 - points_df$X1)^2 + (1 - points_df$X2)^2)
  dist_vertex <- pmin(dist_v1, dist_v2, dist_v3, dist_v4)
  
  # 计算满足条件的比例
  prop_edge <- sum(dist_edge < 0.25) / n_obs
  prop_vertex <- sum(dist_vertex < 0.25) / n_obs
  
  # 返回一个列表包含两个比例
  list(
    prop_near_edge = prop_edge,
    prop_near_vertex = prop_vertex
  )
}

# 使用10000个模拟点进行计算
many_points <- simulate_points(10000)
proportions <- calculate_proportions(many_points)
print(proportions)
```

## 5.

To estimate $\pi$ with a Monte Carlo simulation, we draw the unit circle inside the unit square, the ratio of the area of the circle to the area of the square will be $\pi / 4$. Then shot $K$ arrows at the square, roughly $K * \pi / 4$ should have fallen inside the circle. So if now you shoot $N$ arrows at the square, and $M$ fall inside the circle, you have the following relationship $M = N * \pi / 4$. You can thus compute $\pi$ like so: $\pi = 4 * M / N$. The more arrows $N$ you throw at the square, the better approximation of $\pi$ you'll have.

```{r}
n <- 10000

set.seed(1)
points <- tibble("x" = runif(n), "y" = runif(n))
```

Now, to know if a point is inside the unit circle, we need to check whether $x^2 + y^2 < 1$. Let's add a new column to the points tibble, called `inside` equal to `1` if the point is inside the unit circle and `0` if not:

```{r}
points <- points |> 
    mutate(inside = map2_dbl(.x = x, .y = y, ~ifelse(.x**2 + .y**2 < 1, 1, 0))) |> 
    rowid_to_column("N")
```

a. Compute the estimation of $\pi$ at each row, by computing the cumulative sum of the 1's in the `inside` column and dividing that by the current value of `N` column:

```{r}
# cumsum(inside) 计算落在圆内的点的累积数量 M
# pi_est = 4 * M / N
pi_estimates <- points %>%
  mutate(
    M = cumsum(inside),
    pi_est = 4 * M / N
  )

print("Pi estimates at each step (first 10):")
kable(head(pi_estimates, 10))
```

b. Plot the estimates of $\pi$ against `N`.

```{r}
ggplot(pi_estimates, aes(x = N, y = pi_est)) +
  geom_line(color = "blue", alpha = 0.7) +
  # 添加一条水平红线表示pi的真实值
  geom_hline(yintercept = pi, color = "red", linetype = "dashed") +
  # 设置图表标题和坐标轴标签 (英文)
  labs(
    title = "Monte Carlo Estimation of Pi",
    x = "Number of Throws (N)",
    y = "Estimate of Pi"
  ) +
  theme_minimal()
```

## 6. 
Mortality rates per 100,000 from male suicides for a number of age groups and a number of countries are given in the following data frame. 
```{r}
suicrates <- tibble(Country = c('Canada', 'Israel', 'Japan', 'Austria', 'France', 'Germany',
'Hungary', 'Italy', 'Netherlands', 'Poland', 'Spain', 'Sweden', 'Switzerland', 'UK', 'USA'), 
Age25.34 = c(22,  9, 22, 29, 16, 28, 48,  7,  8, 26,  4, 28, 22, 10, 20), 
Age35.44 = c(27, 19, 19, 40, 25, 35, 65,  8, 11, 29,  7, 41, 34, 13, 22), 
Age45.54 = c(31, 10, 21, 52, 36, 41, 84, 11, 18, 36, 10, 46, 41, 15, 28), 
Age55.64 = c(34, 14, 31, 53, 47, 49, 81, 18, 20, 32, 16, 51, 50, 17, 33), 
Age65.74 = c(24, 27, 49, 69, 56, 52, 107, 27, 28, 28, 22, 35, 51, 22, 37))
```

a. Transform `suicrates` into *long* form.
```{r}
# 使用 pivot_longer 将宽数据转换为长数据
suicrates_long <- suicrates %>%
  pivot_longer(
    cols = -Country,          # 转换除Country外的所有列
    names_to = "AgeGroup",    # 新的列名，存储原列名
    values_to = "Rate"        # 新的列名，存储原单元格值
  )

print("Transformed data in long format (first 6 rows):")
kable(head(suicrates_long))
```

b. Construct side-by-side box plots for the data from different age groups, and comment on what the graphic tells us about the data. 
```{r}
ggplot(suicrates_long, aes(x = AgeGroup, y = Rate, fill = AgeGroup)) +
  geom_boxplot() +
  # 设置图表标题和坐标轴标签 (英文)
  labs(
    title = "Male Suicide Rates by Age Group",
    x = "Age Group",
    y = "Mortality Rate per 100,000"
  ) +
  theme_minimal() +
  # 旋转x轴标签以防重叠
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") # 隐藏图例
```
**评论:** 从箱线图中可以看出，随着年龄的增长，男性自杀率的中位数和分布范围（即四分位距）整体上呈现上升趋势。
`Age65-74`年龄组的自杀率最高，且有一个非常高的异常值，表明在某些国家该年龄段的自杀问题尤为严重。
具体来说，该箱线图揭示了男性自杀率与年龄的显著关系。主要观察如下：
整体趋势：随着年龄组的增长，自杀率的中位数呈现出明显的上升趋势，表明年龄较大的群体面临更高的自杀风险。
数据离散度：年龄越大，箱体（即四分位距）也越宽，这说明不同国家在高年龄组的自杀率差异更大，数据变异性增加。
关键异常值：Age65-74 年龄组不仅拥有最高的中位数，还包含一个极高的异常值，这说明在至少一个国家，该年龄段的自杀率远超其他国家，是一个需要特别关注的公共卫生问题。

## 7. 
Load the `LaborSupply` dataset from the `{Ecdat}` package and answer the following questions:

```{r}
#data(LaborSupply)
LaborSupply <- read_csv("LaborSupply.csv")
# create hour and wage variables
labor <- LaborSupply |> 
  mutate(hour = exp(lnhr), wage = exp(lnwg), .before = kids) |> 
  dplyr::select(-lnhr, -lnwg)
```

a. Compute the average annual hours worked and their standard deviations by year.

```{r}
# 从提供的 "LaborSupply.csv" 文件加载数据
LaborSupply <- read_csv("LaborSupply.csv")

# 创建 hour 和 wage 变量
labor <- LaborSupply %>% 
  mutate(hour = exp(lnhr), wage = exp(lnwg), .before = kids) %>% 
  dplyr::select(-lnhr, -lnwg)

# 按年份分组，然后计算每年的平均工时和标准差
summary_by_year <- labor %>%
  group_by(year) %>%
  summarise(
    Average.Hours = mean(hour),
    Std.Dev.Hours = sd(hour)
  )

print("Average annual hours worked by year:")
kable(summary_by_year)
```

b. What age group worked the most hours in the year 1982?

```{r}
# 筛选1982年的数据，并使用cut()将年龄分组成段
# 然后按年龄组计算平均工时，并排序找到最高值
most_hours_group <- labor %>%
  filter(year == 1982) %>%
  mutate(
    Age.Group = cut(age, breaks = c(20, 30, 40, 50, 60), 
                    right = FALSE)
  ) %>%
  group_by(Age.Group) %>%
  summarise(Average.Hours = mean(hour)) %>%
  arrange(desc(Average.Hours))

print("Average hours by age group in 1982:")
kable(most_hours_group)
print(paste("The age group with the most hours is:", 
            most_hours_group$Age.Group[1]))
```

c. Create a variable, `n_years` that equals the number of years an individual stays in the panel. Is the panel balanced?

```{r}
# 按个体id分组，使用n()计算每个id出现的次数
labor_with_nyears <- labor %>%
  group_by(id) %>%
  mutate(n_years = n()) %>%
  ungroup()

# 检查所有n_years是否都等于最大值
# 如果是，则面板数据是平衡的
is_balanced <- all(labor_with_nyears$n_years == max(labor_with_nyears$n_years))

print(paste("Is the panel balanced?", is_balanced))
print("Distribution of n_years:")
kable(table(labor_with_nyears$n_years))
```

d. Which are the individuals that do not have any kids during the whole period? Create a variable, `no_kids`, that flags these individuals (1 = no kids, 0 = kids)

```{r}
# 按id分组，如果一个id的所有kids加起来等于0，则标记为1
labor_with_nokids <- labor %>%
  group_by(id) %>%
  mutate(no_kids = ifelse(sum(kids) == 0, 1, 0)) %>%
  ungroup()

print("Individuals with no kids (first 6 rows):")
kable(head(filter(labor_with_nokids, no_kids == 1)))
```

e. Using the `no_kids` variable from before compute the average wage, standard deviation and number of observations in each group for the year 1980 (no kids group vs kids group).

```{r}
# 筛选1980年的数据，按no_kids分组
# 计算每组的平均工资、标准差和观测数(n())
summary_1980 <- labor_with_nokids %>%
  filter(year == 1980) %>%
  group_by(no_kids) %>%
  summarise(
    Average.Wage = mean(wage),
    Std.Dev.Wage = sd(wage),
    Observations = n()
  )

print("Summary of wage in 1980 by kids status:")
kable(summary_1980)
```

```{r}
```
