---
title: "Homework 5: Pareto and Kuznets on the Grand Tour"
documentclass: ctexart
output:
  pdf_document:
      latex_engine: xelatex
---

We continue working with the World Top Incomes Database [https://wid.world], and the Pareto distribution, as in the lab.  We also continue to practice working with data frames, manipulating data from one format to another, and writing functions to automate repetitive tasks.

We saw in the lab that if the upper tail of the income distribution followed a perfect Pareto distribution, then
\begin{eqnarray}
\label{eqn:1percent-vs-0.1-percent}
\left(\frac{P99}{P99.9}\right)^{-a+1}  & = & 10\\
\left(\frac{P99.5}{P99.9}\right)^{-a+1} & = & 5\\
\left(\frac{P99}{P99.5}\right)^{-a+1} & = & 2
\label{eqn:1percent-vs-0.5-percent}
\end{eqnarray}
We could estimate the Pareto exponent by solving any one of these equations for $a$; in lab we used
\begin{equation}
a = 1 - \frac{\log{10}}{\log{(P99/P99.9)}} ~,
\label{eqn:exponent-from-quantile-ratio}
\end{equation}

Because of measurement error and sampling noise, we can't find find one value of $a$ which will work for all three equations \eqref{eqn:1percent-vs-0.1-percent}--\eqref{eqn:1percent-vs-0.5-percent}.  Generally, trying to make all three equations come close to balancing gives a better estimate of $a$ than just solving one of them.  (This is analogous to finding the slope and intercept of a regression line by trying to come close to all the points in a scatterplot, and not just running a line through two of them.)

```{r}
library(dplyr)      # 数据处理
library(ggplot2)    # 绘图
library(readr)      # 读取CSV文件
```

1. We estimate $a$ by minimizing
\[
\left(\left(\frac{P99}{P99.9}\right)^{-a+1} - 10\right)^2 + \left(\left(\frac{P99.5}{P99.9}\right)^{-a+1} - 5\right)^2 +  \left(\left(\frac{P99}{P99.5}\right)^{-a+1} - 2\right)^2
\]
Write a function, `percentile_ratio_discrepancies`, which takes as inputs `P99`, `P99.5`, `P99.9` and `a`, and returns the value of the expression above.  Check that when `P99=1e6`, `P99.5=2e6`, `P99.9=1e7` and `a=2`, your function returns `0`.

```{r}
# 编写函数 percentile_ratio_discrepancies
# 计算百分位数比值与理论值的偏差平方和

percentile_ratio_discrepancies <- function(P99, P99.5, P99.9, a) {
  # 输入参数：
  # P99: 99%分位数
  # P99.5: 99.5%分位数
  # P99.9: 99.9%分位数
  # a: 帕累托指数

  # 根据帕累托分布理论，计算三个比值的偏差
  # 第一个等式：(P99/P99.9)^(-a+1) = 10
  term1 <- ((P99 / P99.9)^(-a + 1) - 10)^2
  # 第二个等式：(P99.5/P99.9)^(-a+1) = 5
  term2 <- ((P99.5 / P99.9)^(-a + 1) - 5)^2
  # 第三个等式：(P99/P99.5)^(-a+1) = 2
  term3 <- ((P99 / P99.5)^(-a + 1) - 2)^2
  # 返回偏差平方和
  return(term1 + term2 + term3)
}

# 检验函数：当P99=1e6, P99.5=2e6, P99.9=1e7, a=2时，函数应返回0
test_result_1 <- percentile_ratio_discrepancies(P99 = 1e6, P99.5 = 2e6,
                                                P99.9 = 1e7, a = 2)
cat("问题1检验结果：", test_result_1, "\n")  # 应该输出接近0的值
```

2. Write a function, `exponent.multi_ratios_est`, which takes as inputs `P99`, `P99.5`, `P99.9`, and estimates `a`.  It should minimize your `percentile_ratio_discrepancies` function.  The starting value for the minimization should come from \eqref{eqn:exponent-from-quantile-ratio}.  Check that when `P99=1e6`, `P99.5=2e6` and `P99.9=1e7`, your function returns an `a` of 2.

```{r}
# 编写函数 exponent.multi_ratios_est
# 通过最小化偏差函数来估计帕累托指数a

exponent.multi_ratios_est <- function(P99, P99.5, P99.9) {
  # 输入参数：三个百分位数值
  # 返回：估计的帕累托指数a

  # 使用公式(4)计算初始值作为优化起点
  initial_a <- 1 - log(10) / log(P99 / P99.9)
  # 定义目标函数（要最小化的函数）
  objective_function <- function(a) {
    return(percentile_ratio_discrepancies(P99, P99.5, P99.9, a))
  }

  # 使用optimize函数进行一维优化
  # 搜索区间设定为[0.5, 5]，覆盖常见的帕累托指数范围
  result <- optimize(objective_function, 
                     interval = c(0.5, 5),  # 搜索区间
                     tol = 1e-10)           # 精度要求
  return(result$minimum)  # 返回最优的a值
}

# 检验函数：使用测试数据验证
test_result_2 <- exponent.multi_ratios_est(P99 = 1e6, P99.5 = 2e6, P99.9 = 1e7)
cat("问题2检验结果：估计的a值 =", test_result_2, "\n")  # 应该接近2
```

3. Write a function which uses  `exponent.multi_ratios_est` to estimate $a$ for the US for every year from 1913 to 2012.  (There are many ways you could do thi, including loops.)  Plot the estimates; make sure the labels of the plot are appropriate.

```{r}
# 读取数据
data <- read_csv("data/wtid-report.csv")

# 查看数据结构
head(data)
str(data)

# 筛选美国数据、重命名列并选择需要的列
us_data <- data %>%
  # 重命名列，以匹配后续代码
  rename(
    P99 = `P99 income threshold`,
    P99.5 = `P99.5 income threshold`,
    P99.9 = `P99.9 income threshold`
  ) %>%
  filter(Country == "United States") %>%  # 筛选美国数据
  filter(Year >= 1913 & Year <= 2012) %>%  # 筛选年份范围
  select(Year, P99, P99.5, P99.9) %>%      # 选择需要的百分位数列
  filter(!is.na(P99) & !is.na(P99.5) & !is.na(P99.9))  # 移除缺失值

# 为每年估计Pareto指数
estimate_pareto_by_year <- function(data) {
  # 初始化结果向量
  years <- c()
  pareto_estimates <- c()

  # 遍历每年的数据
  for (year in unique(data$Year)) {
    year_data <- data[data$Year == year, ]
    # 检查是否有必要的百分位数据
    if (all(c("P99", "P99.5", "P99.9") %in% names(year_data)) &&
          !any(is.na(c(year_data$P99, year_data$P99.5, year_data$P99.9)))) {
      # 估计该年的Pareto指数
      a_est <- exponent.multi_ratios_est(
        P99 = year_data$P99,
        P99.5 = year_data$P99.5,
        P99.9 = year_data$P99.9
      )
      years <- c(years, year)
      pareto_estimates <- c(pareto_estimates, a_est)
    }
  }
  return(data.frame(Year = years, Pareto_Exponent = pareto_estimates))
}

# 估计所有年份的Pareto指数
us_pareto_estimates <- estimate_pareto_by_year(us_data)

# 绘制时间序列图
ggplot(us_pareto_estimates, aes(x = Year, y = Pareto_Exponent)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(color = "red", size = 2) +
  labs(
    title = "Pareto Exponent Estimates for United States (1913-2012)",
    # 美国Pareto指数估计(1913-2012)
    x = "Year",  # 年份
    y = "Pareto Exponent (a)",  # Pareto指数(a)
    caption = "Estimated using multi-ratio method"  # 使用多比率方法估计
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

4. Use \eqref{eqn:exponent-from-quantile-ratio} to estimate $a$ for the US for every year.  Make a scatter-plot of these estimates against those from problem 3.  If they are identical or completely independent, something is wrong with at least one part of your code.  Otherwise, can you say anything about how the two estimates compare?

```{r}
# 第4题：使用单一比率公式估计并比较两种方法

# 使用公式 a = 1 - log(10) / log(P99/P99.9) 估计帕累托指数
us_data$single_ratio_a <- 1 - log(10) / log(us_data$P99 / us_data$P99.9)

# 合并两种估计方法的结果
comparison_data <- data.frame(
  Year = us_data$Year,
  Multi_Ratio = us_pareto_estimates$Pareto_Exponent,  # 多比率方法
  Single_Ratio = us_data$single_ratio_a  # 单一比率方法
)

# 创建散点图比较两种方法
plot2 <- ggplot(comparison_data, aes(x = Single_Ratio, y = Multi_Ratio)) +
  geom_point(color = "darkgreen", alpha = 0.7, size = 2) +  # 深绿色半透明点
  geom_abline(intercept = 0, slope = 1, color = "red",
              linetype = "dashed") +  # 45度参考线
  labs(
    title = "Comparison of Pareto Exponent Estimation Methods", 
    # 主标题：帕累托指数估计方法比较
    subtitle = "Multi-ratio vs Single-ratio estimates for US (1913-2012)", 
    # 副标题：美国多比率与单一比率估计比较
    x = "Single Ratio Method Estimate",  # x轴：单一比率方法估计
    y = "Multi-ratio Method Estimate"   # y轴：多比率方法估计
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

print(plot2)  # 显示比较图

# 结果分析和统计总结

# 计算两种方法的相关系数
correlation <- cor(comparison_data$Single_Ratio,
                   comparison_data$Multi_Ratio, use = "complete.obs")
cat("\n两种估计方法的相关系数：", round(correlation, 4), "\n")

# 计算差异的统计量
differences <- comparison_data$Multi_Ratio - comparison_data$Single_Ratio
cat("估计差异的统计摘要：\n")
print(summary(differences))

# 计算均方根误差
rmse <- sqrt(mean(differences^2, na.rm = TRUE))
cat("均方根误差（RMSE）：", round(rmse, 4), "\n")

# 创建时间序列比较图
plot3 <- ggplot(comparison_data) +
  geom_line(aes(x = Year, y = Multi_Ratio, color = "Multi-ratio"),
            linewidth = 1) +  # 多比率方法线条
  geom_line(aes(x = Year, y = Single_Ratio, color = "Single-ratio"),
            linewidth = 1) +  # 单一比率方法线条
  scale_color_manual(
    name = "Method",  # 图例标题：方法
    values = c("Multi-ratio" = "blue", "Single-ratio" = "red")  # 颜色设置
  ) +
  labs(
    title = "Time Series Comparison of Pareto Exponent Estimates",
    # 主标题：帕累托指数估计的时间序列比较
    subtitle = "United States (1913-2012)",  # 副标题：美国
    x = "Year",  # x轴：年份
    y = "Pareto Exponent (a)"  # y轴：帕累托指数
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "bottom"  # 图例位置在底部
  )

print(plot3)  # 显示时间序列比较图

# 输出部分数据查看结果
cat("\n前10年的估计结果对比：\n")
print(head(comparison_data, 10))
```