---
title: "Homework 2"
documentclass: ctexart
output:
  pdf_document:
    latex_engine: xelatex
---
```{r setup, message = F, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(DAAG)
```

The data set calif_penn_2011.csv contains information about the housing stock of California and Pennsylvania, as of 2011.  Information as aggregated into "Census tracts", geographic regions of a few thousand people which are supposed to be fairly homogeneous economically and socially.

1. _Loading and cleaning_
    a. Load the data into a dataframe called `ca_pa`.
    b. How many rows and columns does the dataframe have?
    c. Run this command, and explain, in words, what this does:
```
colSums(apply(ca_pa,c(1,2),is.na))
```
    d. The function `na.omit()` takes a dataframe and returns a new dataframe, omitting any row containing an NA value.  Use it to purge the data set of rows with incomplete data.
    e. How many rows did this eliminate?
    f. Are your answers in (c) and (e) compatible? Explain.

```{r}
ca_pa <- read.csv("data/calif_penn_2011.csv")
dim(ca_pa)
colSums(apply(ca_pa, c(1, 2), is.na))
ca_pa_clean <- na.omit(ca_pa)
nrow(ca_pa) - nrow(ca_pa_clean)
```

* T1答：（a）代码见上；（b）代码见上，11275行和34列；(c)该命令用于计算`ca_pa`中每一列包含的缺失值（NA）的数量；
（d）代码见上；（e）代码见上，670行。

2. _This Very New House_  
    a. The variable `Built_2005_or_later` indicates the percentage of houses in each Census tract built since 2005.  Plot median house prices against this variable.
    b. Make a new plot, or pair of plots, which breaks this out by state.  Note that the state is recorded in the `STATEFP` variable, with California being state 6 and Pennsylvania state 42.

```{r}
# a
plot(ca_pa_clean$Built_2005_or_later, ca_pa_clean$Median_House_Value,
     xlab = "Percentage of houses built since 2005",
     ylab = "Median House Value",
     main = "Median House Value vs. Built Since 2005")
# b
ggplot(ca_pa_clean, aes(x = Built_2005_or_later, y = Median_house_value,
                        color = factor(STATEFP))) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ STATEFP, labeller = as_labeller(c(`6` = "California(6)",
                                                 `42` = "pennsylvania(42)"))) +
  labs(x = "Percentage of houses built since 2005",
       y = "Median House Value",
       color = "State") +
  ggtitle("Median House Value vs. Built Since 2005 by State")
```

3. _Nobody Home_  
The vacancy rate is the fraction of housing units which are not occupied.  The dataframe contains columns giving the total number of housing units for each Census tract, and the number of vacant housing units.  
    a. Add a new column to the dataframe which contains the vacancy rate.  What are the minimum, maximum, mean, and median vacancy rates?
    b. Plot the vacancy rate against median house value.
    c. Plot vacancy rate against median house value separately for California and for Pennsylvania.  Is there a difference?

```{r}
# a
ca_pa_clean$Vacancy_rate <- ca_pa_clean$Vacant_units / ca_pa_clean$Total_units
summary(ca_pa_clean$Vacancy_rate)
# b
ggplot(ca_pa_clean, aes(x = Vacancy_rate, y = Median_house_value)) +
  geom_point(alpha = 0.5) +
  labs(x = "Vacancy Rate", y = "Median House Value") +
  ggtitle("Median Rate vs. Vacancy House Value")
# c
ggplot(ca_pa_clean, aes(x = Vacancy_rate, y = Median_house_value,
                        color = factor(STATEFP))) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ STATEFP, labeller = as_labeller(c(`6` = "California(6)",
                                                 `42` = "pennsylvania(42)"))) +
  labs(x = "Vacancy Rate", y = "Median House Value", color = "State") +
  ggtitle("Median House Value vs. Vacancy Rate by State")
```

* T3答： （a）代码见上,空置率最小为0.00000，最大为0.96531，平均为0.08889，中位数为0.06767；（b）代码见上；
（c）代码见上，注意到有区别，且具体来说：首先，在绝对价格上，加利福尼亚州的房价中位数远高于宾夕法尼亚州。
其次，在数据分布上，加州的房屋市场更为紧俏，绝大多数地区的空置率都紧密集中在25%以下；而宾州的空置率分布则相对更广泛。
在关系趋势方面，两个州都表现出房价中位数和空置率之间的负相关性。这种关系主要体现为：最高价位的房产只存在于低空置率的地区。
随着空置率的增加，房价的上限随之降低，这表明高空置率对维持高昂的房产价值有很强的抑制作用。

4. The column `COUNTYFP` contains a numerical code for counties within each state.  We are interested in Alameda County (county 1 in California), Santa Clara (county 85 in California), and Allegheny County (county 3 in Pennsylvania).
    a. Explain what the block of code at the end of this question is supposed to accomplish, and how it does it.
    b. Give a single line of R which gives the same final answer as the block of code.  Note: there are at least two ways to do this; you just have to find one.
    c. For Alameda, Santa Clara and Allegheny Counties, what were the average percentages of housing built since 2005?
    d. The `cor` function calculates the correlation coefficient between two variables.  What is the correlation between median house value and the percent of housing built since 2005 in (i) the whole data, (ii) all of California, (iii) all of Pennsylvania, (iv) Alameda County, (v) Santa Clara County and (vi) Allegheny County?
    e. Make three plots, showing median house values against median income, for Alameda, Santa Clara, and Allegheny Counties.  (If you can fit the information into one plot, clearly distinguishing the three counties, that's OK too.)
```{r, eval=FALSE}
acca <- c()
for (tract in 1:nrow(ca_pa)) {
  if (ca_pa$STATEFP[tract] == 6) {
    if (ca_pa$COUNTYFP[tract] == 1) {
      acca <- c(acca, tract)
    }
  }
}
accamhv <- c()
for (tract in acca) {
  accamhv <- c(accamhv, ca_pa[tract, 10])
}
median(accamhv)
```

```{r}
# b
ca_pa_clean %>%
  filter(STATEFP == 6, COUNTYFP == 1) %>%
  summarise(Median_house_value = median(Median_house_value))
median(ca_pa_clean$Median_house_value[ca_pa_clean$STATEFP == 6 &
                                        ca_pa_clean$COUNTYFP == 1])
# c
ca_pa_clean %>%
  filter((STATEFP == 6 & COUNTYFP %in% c(1, 85)) |
           (STATEFP == 42 & COUNTYFP == 3)) %>%
  group_by(STATEFP, COUNTYFP) %>%
  summarise(avg_built_since_2005 = mean(Built_2005_or_later))
# d
cor(ca_pa_clean$Median_house_value, ca_pa_clean$Built_2005_or_later,
    use = "complete.obs")
ca_ca <- filter(ca_pa_clean, STATEFP == 6)
cor(ca_ca$Median_house_value, ca_ca$Built_2005_or_later,
    use = "complete.obs")
pa_pa <- filter(ca_pa_clean, STATEFP == 42)
cor(pa_pa$Median_house_value, pa_pa$Built_2005_or_later,
    use = "complete.obs")
acca <- filter(ca_pa_clean, STATEFP == 6, COUNTYFP == 1)
cor(acca$Median_house_value, acca$Built_2005_or_later,
    use = "complete.obs")
accamhv <- filter(ca_pa_clean, STATEFP == 6, COUNTYFP == 85)
cor(accamhv$Median_house_value, accamhv$Built_2005_or_later,
    use = "complete.obs")
allegheny <- filter(ca_pa_clean, STATEFP == 42, COUNTYFP == 3)
cor(allegheny$Median_house_value, allegheny$Built_2005_or_later,
    use = "complete.obs")
# e
three_counties <- ca_pa_clean %>%
  filter((STATEFP == 6 & COUNTYFP %in% c(1, 85)) |
           (STATEFP == 42 & COUNTYFP == 3)) %>%
  mutate(CountyName = case_when(
    STATEFP == 6 & COUNTYFP == 1 ~ "Alameda,CA",
    STATEFP == 6 & COUNTYFP == 85 ~ "Santa Clara,CA",
    STATEFP == 42 & COUNTYFP == 3 ~ "Allegheny,PA",
    TRUE ~ "Other"
  ))
ggplot(three_counties, aes(x = Median_household_income, y = Median_house_value,
                           color = CountyName)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Median House Value vs. 
              Median Household Income in Three Counties",
       x = "Median Income", y = "Median House Value",
       color = "County") +
  theme_minimal() +
  scale_color_manual(values = c("Alameda,CA" = "blue",
                                "Santa Clara,CA" = "red",
                                "Allegheny,PA" = "green"))
```
* T4答：（a）该代码块的目的是计算加利福尼亚州阿拉米达县（COUNTYFP=1）中房屋建造于2005年或之后的房屋的中位数房价。
它通过遍历每个地块，检查其州和县代码，并将符合条件的地块的房价存储在`accamhv`向量中，最后计算并返回该向量的中位数；
（b）代码见上，也得到Median_house_value为474050；（c）代码见上，平均百分比依次为2.82,3.20,1.47；
（d）代码见上；（e）代码见上。

MB.Ch1.11. Run the following code:
```{r}
gender <- factor(c(rep("female", 91), rep("male", 92)))
table(gender)
gender <- factor(gender, levels = c("male", "female"))
table(gender)
gender <- factor(gender, levels = c("Male", "female"))
# Note the mistake: "Male" should be "male"
table(gender)
table(gender, exclude = NULL)
rm(gender)  # Remove gender
```
Explain the output from the successive uses of table().

* Ch1.11答：*
*1.第一次使用:首次创建因子时，默认按照字母顺序排列，因此“female”在“male”之前。输出为`female` 91, `male` 92。*
*2.第二次使用:通过指定levels参数（`levels=c("Male", "female")`），将因子水平更改为`male`和`female`，
因此输出顺序也相应更改。输出为`male` 92, `female` 91。*
*3.第三次使用:指定levels参数（`levels=c("Male", "female")`），将因子水平更改为`Male`和`female`，
由于“Male”与“male”不匹配，因此“Male”被视为一个新的水平，从而输出为`Male` 0, `female` 91。*
*4.第四次使用:通过设置`exclude = NULL`，将所有因子水平都包含在输出中，即使它们没有观察值。
于是`table()` 函数也要统计 `NA` 值。因此输出为`Male` 0,`female` 91, `<NA>` 92 。
这92次其实对应于那些与新水平 `"Male"` 不匹配的原始 `"male"` 值。
     
MB.Ch1.12. Write a function that calculates the proportion of values in a vector x that exceed some value cutoff.

(a) Use the sequence of numbers 1, 2, . . . , 100 to check that this function gives the result that is expected.
```{r}
proportion_exceeding <- function(x, cutoff) {
  num_exceeding <- sum(x > cutoff)
  proportion <- num_exceeding / length(x)
  return(proportion)
}
# 测试函数
test_vector <- 1:100
proportion_exceeding(test_vector, 90) # 应返回0.1
proportion_exceeding(test_vector, 50) # 应返回0.5
```

(b) Obtain the vector ex01.36 from the Devore6 (or Devore7) package. These data give the times required for individuals to escape from an oil platform during a drill. Use dotplot() to show the distribution of times. Calculate the proportion of escape times that exceed 7 minutes.
```{r}
# 请预先安装Devore7包
if (!requireNamespace("Devore7", quietly = TRUE)) {
  install.packages("Devore7")
}
library(Devore7)
data(ex01.36)
dotplot(~ C1, data = ex01.36, main = "Escape Times from Oil Platform",
        xlab = "Time (seconds)", ylab = "Frequency")
proportion_exceeding(ex01.36$C1, 420) # 超过7分钟（420秒）的比例
```

MB.Ch1.18. The Rabbit data frame in the MASS library contains blood pressure change measurements on five rabbits (labeled as R1, R2, . . . ,R5) under various control and treatment conditions. Read the help file for more information. Use the unstack() function (three times) to convert Rabbit to the following form:

Treatment Dose   R1   R2   R3   R4   R5 

1 Control 6.25  0.50 1.00 0.75 1.25 1.5 

2 Control 12.50 4.50 1.25 3.00 1.50 1.5 

....
```{r}
library(MASS)
data(Rabbit)

# 第一次 unstack：按兔子ID将血压变化分开
unstacked1 <- unstack(Rabbit, BPchange ~ Animal)
# 第二次 unstack：按兔子ID将剂量分开
unstacked2 <- unstack(Rabbit, Dose ~ Animal)
# 第三次 unstack：按兔子ID将治疗方法分开
unstacked3 <- unstack(Rabbit, Treatment ~ Animal)

# 合并结果
final_df <- data.frame(
  Treatment = unstacked3$R1,
  Dose = unstacked2$R1,
  R1 = unstacked1$R1,
  R2 = unstacked1$R2,
  R3 = unstacked1$R3,
  R4 = unstacked1$R4,
  R5 = unstacked1$R5
)

print(final_df)
```

